---
title: "startCh7"
author: "Jin Chen"
date: "7/26/2020"
output: html_document
---

```{r}
library("causact")
data("delivDF")
delivDF

```

```{r 7.2 Lateness Calculation}
library(lubridate)

shipDF = delivDF
shipDF$plannedShipDate = mdy(shipDF$plannedShipDate) #convert character to Date
shipDF$actualShipDate = mdy(shipDF$actualShipDate)   #convert character to Date
shipDF$actualShipDate[1] - shipDF$plannedShipDate[1]
```
As an analyst, it is your job to form an opinion and validate that opinion with your stakeholder about how you plan to translate real-world concerns into mathematical constructs. Do not immediately fire off an email everytime you have a question

```{r}
library(dplyr)

shipDateDF = shipDF %>% 
  select(shipID, plannedShipDate, actualShipDate) %>% 
  distinct() ## get unique rows to avoid double counting
shipDateDF

#Add a column to capture lateness
shipDateDF = shipDateDF %>% 
  mutate(lateFlag = 
           ifelse(actualShipDate > plannedShipDate, #ifelse returns TRUE if condition is TRUE and FALSE if condition is FALSE
                  TRUE,
                  FALSE))
shipDateDF

#now we count the number of TRUE and FALSE
shipDateDF %>% 
  group_by(lateFlag) %>% 
  summarize(countLate = n()) %>% 
  mutate(proportion = countLate / sum(countLate))
```

```{r 7.3 Bringing in Product Category Information}
library(causact)
data("prodLineDF")
prodLineDF

catDF = shipDF %>% 
  left_join(prodLineDF) %>% 
  filter(!is.na(prodCategory)) #filter out any NA
```

**Does service level (measured by on-time shipments) vary across product categories?**
```{r 7.4 Answering the CEO's Questions}
catDF %>% 
  select(shipID, plannedShipDate, actualShipDate, prodCategory) %>% 
  distinct() %>%  #one row per shipment/prodCategory
  mutate(lateFlag = 
           ifelse(actualShipDate > plannedShipDate,
                  TRUE,
                  FALSE)) %>% 
  group_by(prodCategory, lateFlag) %>% 
  summarize(countLate = n()) %>% 
  mutate(proportion = countLate / sum(countLate)) %>% 
  arrange(lateFlag, proportion)

```
From the above analysis, we find that there does seem to be discrepancies between on-time shipments by product category. Machines has the most late shipments (19%), SpareParts(15.8%) is next, and the remaining two, Liquids (8.1%) and Marketables (8.1%) have similar performance.

**How often do orders include products from more than one product category?**
```{r}
numCatDF = catDF %>% 
  select(shipID, plannedShipDate, actualShipDate, prodCategory) %>% 
  distinct() %>%  #unique rows only
  group_by(shipID) %>% 
  summarize(numCategories = n())

#print out summary
numCatDF %>% 
  group_by(numCategories) %>% 
  summarize(numShipID = n()) %>% 
  mutate(percentOfShipments =
           numShipID/sum(numShipID))

```
The answer to this second question is about 22.5% of orders contain more than one product category. So, in conclusion, 22.5% of orders have more than one product category on them and yes, it does seem that the product categories are managed differently.


